---
title: "Compare_models"
author: "Sarah Roberts"
date: "1/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#setup 
```{r, echo = F, eval = F}
install.packages('gjam_2.3.tar.gz',repos=NULL, type='source')
library(gjam) #it has been working with 2.3.2 but the predictions are weird 

message('loading local clark files...')
clark <- 'gjamTimeExample-master/clarkFiles/'
files <- list.files(clark)
cat('   ')
message(paste(files, collapse=", "))
for (f in files) {
  path <- file.path(clark, f)
  if (grepl('.R', toupper(f))) source(path)
  else Rcpp::sourceCpp(path)
}

library(MASS)
library(dplyr)
library(reshape2)
library(ggplot2)
```

#RMSE 
I ran I bunch of models in annual_seasonal.rmd and annual_seasonal_morealpha.rmd and I want to compare them based on individual species RMSE and whole model RMSE for the entire community. 

##more interactions from literature 
```{r}
load("season/XRA_1/output.rdata")
load("season/XRA_2/output.rdata")
load("season/XRA_3/output.rdata")
load("season/XRA_4/output.rdata")
load("season/XRA_5/output.rdata")
seasonal_more_specs_XRA_1 <- XRA_1
seasonal_more_specs_XRA_2 <- XRA_2
seasonal_more_specs_XRA_3 <- XRA_3
seasonal_more_specs_XRA_4 <- XRA_4
seasonal_more_specs_XRA_5 <- XRA_5
#didn't run: seasonal_more_specs_XRA_6
rm(XRA_1, XRA_2, XRA_3, XRA_4, XRA_5)
load("annual/XRA_1/output.rdata")
load("annual/XRA_2/output.rdata")
load("annual/XRA_3/output.rdata")
load("annual/XRA_4/output.rdata")
annual_more_specs_XRA_1 <- XRA_1
annual_more_specs_XRA_2 <- XRA_2
annual_more_specs_XRA_3 <- XRA_3
annual_more_specs_XRA_4 <- XRA_4
rm(XRA_1, XRA_2, XRA_3, XRA_4)

```

get RMSE for all species and total  annual 
```{r}
RMSE_tot_annual_more_specs_XRA_1 <- annual_more_specs_XRA_1[["fit"]][["rmspeAll"]]
RMSE_specs_annual_more_specs_XRA_1 <- as.data.frame(annual_more_specs_XRA_1[["fit"]][["rmspeBySpec"]])
RMSE_specs_annual_more_specs_XRA_1$specs <- rownames(RMSE_specs_annual_more_specs_XRA_1)

RMSE_tot_annual_more_specs_XRA_2 <- annual_more_specs_XRA_2[["fit"]][["rmspeAll"]]
RMSE_specs_annual_more_specs_XRA_2 <- as.data.frame(annual_more_specs_XRA_2[["fit"]][["rmspeBySpec"]])
RMSE_specs_annual_more_specs_XRA_2$specs <- rownames(RMSE_specs_annual_more_specs_XRA_2)

RMSE_tot_annual_more_specs_XRA_3 <- annual_more_specs_XRA_3[["fit"]][["rmspeAll"]]
RMSE_specs_annual_more_specs_XRA_3 <- as.data.frame(annual_more_specs_XRA_3[["fit"]][["rmspeBySpec"]])
RMSE_specs_annual_more_specs_XRA_3$specs <- rownames(RMSE_specs_annual_more_specs_XRA_3)

RMSE_tot_annual_more_specs_XRA_4 <- annual_more_specs_XRA_4[["fit"]][["rmspeAll"]]
RMSE_specs_annual_more_specs_XRA_4 <- as.data.frame(annual_more_specs_XRA_4[["fit"]][["rmspeBySpec"]])
RMSE_specs_annual_more_specs_XRA_4$specs <- rownames(RMSE_specs_annual_more_specs_XRA_4)



joined <- left_join(RMSE_specs_annual_more_specs_XRA_4, RMSE_specs_annual_more_specs_XRA_3, by = c("specs"))

joined <- left_join(joined, RMSE_specs_annual_more_specs_XRA_2, by = c("specs"))
joined <- left_join(joined, RMSE_specs_annual_more_specs_XRA_1, by = c("specs"))

rownames(joined) <- joined$specs

colnames(joined) <- c("lit_annual_most_specs", "specs", "lit_annual_mostcommon+7", "lit_annual_mostcommon+3", "lit_annual_mostcommon")

dat.m <- melt(joined,id.vars='specs', measure.vars=c("lit_annual_most_specs", "lit_annual_mostcommon+7", "lit_annual_mostcommon+3", "lit_annual_mostcommon"))

ggplot(data=dat.m, aes(x=variable, y=value)) + geom_boxplot() 
#interesting results here 

```


```{r}
RMSE_tot_seasonal_more_specs_XRA_1 <- seasonal_more_specs_XRA_1[["fit"]][["rmspeAll"]]
RMSE_specs_seasonal_more_specs_XRA_1 <- as.data.frame(seasonal_more_specs_XRA_1[["fit"]][["rmspeBySpec"]])
RMSE_specs_seasonal_more_specs_XRA_1$specs <- rownames(RMSE_specs_seasonal_more_specs_XRA_1)

RMSE_tot_seasonal_more_specs_XRA_2 <- seasonal_more_specs_XRA_2[["fit"]][["rmspeAll"]]
RMSE_specs_seasonal_more_specs_XRA_2 <- as.data.frame(seasonal_more_specs_XRA_2[["fit"]][["rmspeBySpec"]])
RMSE_specs_seasonal_more_specs_XRA_2$specs <- rownames(RMSE_specs_seasonal_more_specs_XRA_2)

RMSE_tot_seasonal_more_specs_XRA_3 <- seasonal_more_specs_XRA_3[["fit"]][["rmspeAll"]]
RMSE_specs_seasonal_more_specs_XRA_3 <- as.data.frame(seasonal_more_specs_XRA_3[["fit"]][["rmspeBySpec"]])
RMSE_specs_seasonal_more_specs_XRA_3$specs <- rownames(RMSE_specs_seasonal_more_specs_XRA_3)

RMSE_tot_seasonal_more_specs_XRA_4 <- seasonal_more_specs_XRA_4[["fit"]][["rmspeAll"]]
RMSE_specs_seasonal_more_specs_XRA_4 <- as.data.frame(seasonal_more_specs_XRA_4[["fit"]][["rmspeBySpec"]])
RMSE_specs_seasonal_more_specs_XRA_4$specs <- rownames(RMSE_specs_seasonal_more_specs_XRA_4)

RMSE_tot_seasonal_more_specs_XRA_5 <- seasonal_more_specs_XRA_5[["fit"]][["rmspeAll"]]
RMSE_specs_seasonal_more_specs_XRA_5 <- as.data.frame(seasonal_more_specs_XRA_5[["fit"]][["rmspeBySpec"]])
RMSE_specs_seasonal_more_specs_XRA_5$specs <- rownames(RMSE_specs_seasonal_more_specs_XRA_5)


joined_2 <- left_join(RMSE_specs_seasonal_more_specs_XRA_5, RMSE_specs_seasonal_more_specs_XRA_4, by = c("specs"))

joined_2 <- left_join(joined_2, RMSE_specs_seasonal_more_specs_XRA_3, by = c("specs"))
joined_2 <- left_join(joined_2, RMSE_specs_seasonal_more_specs_XRA_2, by = c("specs"))
joined_2 <- left_join(joined_2, RMSE_specs_seasonal_more_specs_XRA_1, by = c("specs"))

rownames(joined_2) <- joined_2$specs

colnames(joined_2) <- c("lit_seasonal_common+3_seasonbeta", "specs", "lit_seasonal_common+3", "lit_seasonal_common_seasonrho", "lit_seasonal_common_seasonbeta", "lit_seasonal_common")

joined_full <- left_join(joined, joined_2, by = c("specs"))

dat.m <- melt(joined_full,id.vars='specs', measure.vars=c("lit_annual_most_specs", "lit_annual_mostcommon+7", "lit_annual_mostcommon+3", "lit_annual_mostcommon", "lit_seasonal_common+3_seasonbeta", "lit_seasonal_common+3", "lit_seasonal_common_seasonrho", "lit_seasonal_common_seasonbeta", "lit_seasonal_common"))

ggplot(data=dat.m, aes(x=variable, y=value)) + geom_boxplot() 
#interesting results here 

```





##just interactions from gut 
```{r}
load("season2/XRA_1/output.rdata")
load("season2/XRA_2/output.rdata")
load("season2/XRA_3/output.rdata")
load("season2/XRA_4/output.rdata")
load("season2/XRA_5/output.rdata")
seasonal_more_specs_XRA_1 <- XRA_1
seasonal_more_specs_XRA_2 <- XRA_2
seasonal_more_specs_XRA_3 <- XRA_3
seasonal_more_specs_XRA_4 <- XRA_4
seasonal_more_specs_XRA_5 <- XRA_5
#didn't run: seasonal_more_specs_XRA_6
rm(XRA_1, XRA_2, XRA_3, XRA_4, XRA_5)
load("annual2/XRA_1/output.rdata")
load("annual2/XRA_2/output.rdata")
load("annual2/XRA_3/output.rdata")
load("annual2/XRA_4/output.rdata")
annual_more_specs_XRA_1 <- XRA_1
annual_more_specs_XRA_2 <- XRA_2
annual_more_specs_XRA_3 <- XRA_3
annual_more_specs_XRA_4 <- XRA_4
rm(XRA_1, XRA_2, XRA_3, XRA_4)

```

get RMSE for all species and total  annual 
```{r}
RMSE_tot_annual_more_specs_XRA_1 <- annual_more_specs_XRA_1[["fit"]][["rmspeAll"]]
RMSE_specs_annual_more_specs_XRA_1 <- as.data.frame(annual_more_specs_XRA_1[["fit"]][["rmspeBySpec"]])
RMSE_specs_annual_more_specs_XRA_1$specs <- rownames(RMSE_specs_annual_more_specs_XRA_1)

RMSE_tot_annual_more_specs_XRA_2 <- annual_more_specs_XRA_2[["fit"]][["rmspeAll"]]
RMSE_specs_annual_more_specs_XRA_2 <- as.data.frame(annual_more_specs_XRA_2[["fit"]][["rmspeBySpec"]])
RMSE_specs_annual_more_specs_XRA_2$specs <- rownames(RMSE_specs_annual_more_specs_XRA_2)

RMSE_tot_annual_more_specs_XRA_3 <- annual_more_specs_XRA_3[["fit"]][["rmspeAll"]]
RMSE_specs_annual_more_specs_XRA_3 <- as.data.frame(annual_more_specs_XRA_3[["fit"]][["rmspeBySpec"]])
RMSE_specs_annual_more_specs_XRA_3$specs <- rownames(RMSE_specs_annual_more_specs_XRA_3)

RMSE_tot_annual_more_specs_XRA_4 <- annual_more_specs_XRA_4[["fit"]][["rmspeAll"]]
RMSE_specs_annual_more_specs_XRA_4 <- as.data.frame(annual_more_specs_XRA_4[["fit"]][["rmspeBySpec"]])
RMSE_specs_annual_more_specs_XRA_4$specs <- rownames(RMSE_specs_annual_more_specs_XRA_4)



joined <- left_join(RMSE_specs_annual_more_specs_XRA_4, RMSE_specs_annual_more_specs_XRA_3, by = c("specs"))

joined <- left_join(joined, RMSE_specs_annual_more_specs_XRA_2, by = c("specs"))
joined <- left_join(joined, RMSE_specs_annual_more_specs_XRA_1, by = c("specs"))

rownames(joined) <- joined$specs

colnames(joined) <- c("gut_annual_most_specs", "specs", "gut_annual_mostcommon+7", "gut_annual_mostcommon+3", "gut_annual_mostcommon")

dat.m <- melt(joined,id.vars='specs', measure.vars=c("gut_annual_most_specs", "gut_annual_mostcommon+7", "gut_annual_mostcommon+3", "gut_annual_mostcommon"))

ggplot(data=dat.m, aes(x=variable, y=value)) + geom_boxplot() 
#interesting results here 

```


```{r}
RMSE_tot_seasonal_more_specs_XRA_1 <- seasonal_more_specs_XRA_1[["fit"]][["rmspeAll"]]
RMSE_specs_seasonal_more_specs_XRA_1 <- as.data.frame(seasonal_more_specs_XRA_1[["fit"]][["rmspeBySpec"]])
RMSE_specs_seasonal_more_specs_XRA_1$specs <- rownames(RMSE_specs_seasonal_more_specs_XRA_1)

RMSE_tot_seasonal_more_specs_XRA_2 <- seasonal_more_specs_XRA_2[["fit"]][["rmspeAll"]]
RMSE_specs_seasonal_more_specs_XRA_2 <- as.data.frame(seasonal_more_specs_XRA_2[["fit"]][["rmspeBySpec"]])
RMSE_specs_seasonal_more_specs_XRA_2$specs <- rownames(RMSE_specs_seasonal_more_specs_XRA_2)

RMSE_tot_seasonal_more_specs_XRA_3 <- seasonal_more_specs_XRA_3[["fit"]][["rmspeAll"]]
RMSE_specs_seasonal_more_specs_XRA_3 <- as.data.frame(seasonal_more_specs_XRA_3[["fit"]][["rmspeBySpec"]])
RMSE_specs_seasonal_more_specs_XRA_3$specs <- rownames(RMSE_specs_seasonal_more_specs_XRA_3)

RMSE_tot_seasonal_more_specs_XRA_4 <- seasonal_more_specs_XRA_4[["fit"]][["rmspeAll"]]
RMSE_specs_seasonal_more_specs_XRA_4 <- as.data.frame(seasonal_more_specs_XRA_4[["fit"]][["rmspeBySpec"]])
RMSE_specs_seasonal_more_specs_XRA_4$specs <- rownames(RMSE_specs_seasonal_more_specs_XRA_4)

RMSE_tot_seasonal_more_specs_XRA_5 <- seasonal_more_specs_XRA_5[["fit"]][["rmspeAll"]]
RMSE_specs_seasonal_more_specs_XRA_5 <- as.data.frame(seasonal_more_specs_XRA_5[["fit"]][["rmspeBySpec"]])
RMSE_specs_seasonal_more_specs_XRA_5$specs <- rownames(RMSE_specs_seasonal_more_specs_XRA_5)


joined_2 <- left_join(RMSE_specs_seasonal_more_specs_XRA_5, RMSE_specs_seasonal_more_specs_XRA_4, by = c("specs"))

joined_2 <- left_join(joined_2, RMSE_specs_seasonal_more_specs_XRA_3, by = c("specs"))
joined_2 <- left_join(joined_2, RMSE_specs_seasonal_more_specs_XRA_2, by = c("specs"))
joined_2 <- left_join(joined_2, RMSE_specs_seasonal_more_specs_XRA_1, by = c("specs"))

rownames(joined_2) <- joined_2$specs

colnames(joined_2) <- c("gut_seasonal_common+3_seasonbeta", "specs", "gut_seasonal_common+3", "gut_seasonal_common_seasonrho", "gut_seasonal_common_seasonbeta", "gut_seasonal_common")

joined_fuller <- left_join(joined, joined_2, by = c("specs"))

colnames(joined_fuller) <- c("gut_annual_most_specs", "specs", "gut_annual_mostcommon+7", "gut_annual_mostcommon+3", "gut_annual_mostcommon", "gut_seasonal_common+3_seasonbeta", "gut_seasonal_common+3", "gut_seasonal_common_seasonrho", "gut_seasonal_common_seasonbeta", "gut_seasonal_common")

dat.m <- melt(joined_fuller,id.vars='specs', measure.vars=c("gut_annual_most_specs", "gut_annual_mostcommon+7", "gut_annual_mostcommon+3", "gut_annual_mostcommon", "gut_seasonal_common+3_seasonbeta", "gut_seasonal_common+3", "gut_seasonal_common_seasonrho", "gut_seasonal_common_seasonbeta", "gut_seasonal_common"))

ggplot(data=dat.m, aes(x=variable, y=value)) + geom_boxplot() 
#interesting results here 

```




##all 

```{r}
joined_fullest <- left_join(joined_fuller, joined_full, by = c("specs"))

dat.m <- melt(joined_fullest,id.vars='specs', measure.vars=c("gut_annual_most_specs", "gut_annual_mostcommon+7", "gut_annual_mostcommon+3", "gut_annual_mostcommon", "gut_seasonal_common+3_seasonbeta", "gut_seasonal_common+3", "gut_seasonal_common_seasonrho", "gut_seasonal_common_seasonbeta", "gut_seasonal_common","lit_annual_most_specs", "lit_annual_mostcommon+7", "lit_annual_mostcommon+3", "lit_annual_mostcommon", "lit_seasonal_common+3_seasonbeta", "lit_seasonal_common+3", "lit_seasonal_common_seasonrho", "lit_seasonal_common_seasonbeta", "lit_seasonal_common"))
```

```{r}
ggplot(data=dat.m, aes(x=variable, y=value)) + geom_boxplot() + theme(axis.text.x = element_text(angle = 90))
#interesting results here 

ggplot(data=dat.m, aes(x=variable, y=value)) + geom_boxplot(outlier.size=-1) + scale_y_continuous(limits = c(0, 200)) + theme(axis.text.x = element_text(angle = 90))
```

#DI and DD 
compare how much variation is explained by DI and DD with different models 

```{r}
load("season/XRA_1/output.rdata")
load("season/XRA_2/output.rdata")
load("season/XRA_3/output.rdata")
load("season/XRA_4/output.rdata")
load("season/XRA_5/output.rdata")
lit_season_XRA_1 <- XRA_1
lit_season_XRA_2 <- XRA_2
lit_season_XRA_3 <- XRA_3
lit_season_XRA_4 <- XRA_4
lit_season_XRA_5 <- XRA_5
#didn't run: seasonal_more_specs_XRA_6
rm(XRA_1, XRA_2, XRA_3, XRA_4, XRA_5)
load("annual/XRA_1/output.rdata")
load("annual/XRA_2/output.rdata")
load("annual/XRA_3/output.rdata")
load("annual/XRA_4/output.rdata")
lit_annual_XRA_1 <- XRA_1
lit_annual_XRA_2 <- XRA_2
lit_annual_XRA_3 <- XRA_3
lit_annual_XRA_4 <- XRA_4
rm(XRA_1, XRA_2, XRA_3, XRA_4)

load("season2/XRA_1/output.rdata")
load("season2/XRA_2/output.rdata")
load("season2/XRA_3/output.rdata")
load("season2/XRA_4/output.rdata")
load("season2/XRA_5/output.rdata")
gut_season_XRA_1 <- XRA_1
gut_season_XRA_2 <- XRA_2
gut_season_XRA_3 <- XRA_3
gut_season_XRA_4 <- XRA_4
gut_season_XRA_5 <- XRA_5
#didn't run: seasonal_more_specs_XRA_6
rm(XRA_1, XRA_2, XRA_3, XRA_4, XRA_5)
load("annual2/XRA_1/output.rdata")
load("annual2/XRA_2/output.rdata")
load("annual2/XRA_3/output.rdata")
load("annual2/XRA_4/output.rdata")
gut_annual_XRA_1 <- XRA_1
gut_annual_XRA_2 <- XRA_2
gut_annual_XRA_3 <- XRA_3
gut_annual_XRA_4 <- XRA_4
rm(XRA_1, XRA_2, XRA_3, XRA_4)
```


##try in a loop 

```{r}

test <- lit_annual_XRA_4
testy <- test[["parameters"]][["sigMu"]]
specs <- rownames(testy)

outputs <- list(gut_annual_XRA_1, gut_annual_XRA_2, gut_annual_XRA_3, gut_annual_XRA_4, gut_season_XRA_1, gut_season_XRA_2, gut_season_XRA_3, gut_season_XRA_4, gut_season_XRA_5, lit_annual_XRA_1, lit_annual_XRA_2, lit_annual_XRA_3, lit_annual_XRA_4, lit_season_XRA_1, lit_season_XRA_2, lit_season_XRA_3, lit_season_XRA_4, lit_season_XRA_5)

outnames <- c("gut_annual_XRA_1", "gut_annual_XRA_2", "gut_annual_XRA_3", "gut_annual_XRA_4", "gut_season_XRA_1", "gut_season_XRA_2", "gut_season_XRA_3", "gut_season_XRA_4", "gut_season_XRA_5", "lit_annual_XRA_1", "lit_annual_XRA_2", "lit_annual_XRA_3", "lit_annual_XRA_4", "lit_season_XRA_1", "lit_season_XRA_2", "lit_season_XRA_3", "lit_season_XRA_4", "lit_season_XRA_5")

    
tot_sprop <- data.frame(matrix(nrow=length(specs), ncol=1))
colnames(tot_sprop) <- "specs"
tot_sprop$specs <- specs


for (i in 1:length(outputs)) {
out <- outputs[i]
sensMu <- sensSe <- numeric(0)
    acol <- colF(3)
    scol <- character(0)
sensSe <- out[[1]][["parameters"]][["sensSe"]]
sigMu <- out[[1]][["parameters"]][["sigMu"]]
sensAlpha <- out[[1]][["parameters"]][["sensAlpha"]]
sensRho <- out[[1]][["parameters"]][["sensRho"]]
sensBeta <- out[[1]][["parameters"]][["sensBeta"]]

sensMu <- cbind(sensMu, sensAlpha[,1])
sensSe <- cbind(sensSe, sensAlpha[,1])
colnames(sensMu)[ncol(sensMu)] <- colnames(sensSe)[ncol(sensMu)] <- 'alpha'
scol <- c(scol, acol[1])

sensMu <- cbind(sensMu, sensRho[,1])
sensSe <- cbind(sensSe, sensRho[,1])
colnames(sensMu)[ncol(sensMu)] <- colnames(sensSe)[ncol(sensMu)] <- 'rho'
scol <- c(scol, acol[2])

sensMu <- cbind(sensMu, sensBeta[,1])
sensSe <- cbind(sensSe, sensBeta[,1])
colnames(sensMu)[ncol(sensMu)] <- colnames(sensSe)[ncol(sensMu)] <- 'beta'
scol <- c(scol, acol[3])

 
      
      nc <- ncol(sensMu)
 
      # total variance
      ord <- order(sensMu[,1], decreasing = T)
    #  scol <- colF(ncol(sensMu))
        
        sigma <- sqrt( diag( sigMu )) #sensAlpha, sensRho, sensBeta on sd scale
        sens  <- cbind(sensMu, sigma)
        
        sens <- sens^2
        
        sprop <- sweep( sens, 1, rowSums(sens), '/')
        sprop <- as.data.frame(sprop)
        sensMu <- as.data.frame(sensMu)
sprop$specs <- rownames(sprop)
sensMu$spec <- rownames(sensMu)
names <- outnames[i]
colnames(sprop) <- paste(names, (colnames(sprop)), sep = "_")
colnames(sprop)[5] <- c("specs")
tot_sprop <- left_join(tot_sprop,sprop, by = c("specs"))

}
    
#sensMu is on the sensitivity scale 
#sprop is on the sd scale 
#I think i need to separate into rho, beta, and alpha proportion of total variance data frames with columns as the model run. Or maybe not im not sure. First i won't separate them. Then I can make a long data frame and seaparte into different columns if I want to. 

```

make that a long table not a wide one 

```{r}
melth_test <- melt(tot_sprop,id.vars='specs', measure.vars=colnames(tot_sprop[2:ncol(tot_sprop)]))

long_dat <- melth_test %>% separate(variable, into = c("model", "stat"), sep="_(?=[^_]+$)")
```

