---
title: "Fishing_pressure"
author: "Sarah Roberts"
date: "1/28/2021"
output: html_document
---
run utilities in gjamTimeExample-master. May need to install gjam and run gjam functions as well. but try just running utilities. Note, its best to go in to the folder, run utilities, then open gjamTime.r and run the first few lines until source(utlities.R)

```{r, echo = F, eval = F}
install.packages('gjam_2.3.tar.gz',repos=NULL, type='source')
library(gjam) #it has been working with 2.3.2 but the predictions are weird 

message('loading local clark files...')
clark <- 'gjamTimeExample-master/clarkFiles/'
files <- list.files(clark)
cat('   ')
message(paste(files, collapse=", "))
for (f in files) {
  path <- file.path(clark, f)
  if (grepl('.R', toupper(f))) source(path)
  else Rcpp::sourceCpp(path)
}

library(MASS)
library(dplyr)
library(tidyverse)
```

#Fishing pressure 
We need to do this now because the number of species changes. So here, I am adding a pre-existing model that I know has the number of species I want. That is the seasonal most common + 7 common prey species. I am going to use only species interactions from the gut content for this test. 


I am going to try and add a fishing mortality covariate based on the species, year, and area collected. The area is going to be split into Mid Atlantic (including NC) and New England. I am going to add recreational and commercial fishing together. 

https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4488883/#pone.0129883.s004

1 Combine recreational and commercial 
	Group by year, state and spp and summarize (sum) pounds .
2. Group into geographic units - based on TNC 
	add column with if else statement 
	group by that column added 
3. Add year column (lagged year) 

4. Select out species that we are using 

5. Group into pelagic, demersal, and benthic fishing mortalities 
	add column based on sticky fish paper 
	group by that column 

6 Add 3 columns (Fdem, Fpel, Fben) into xdata based on year, and geographic unit. 

##Load Data 
```{r}
landings <- read.csv("landings.csv")
landings$Pounds <- as.numeric(gsub(",", "", landings$Pounds))
```


##Group recreation and commercial into one   
```{r}
landings <- landings %>% group_by(Year, State, NMFS.Name) %>% summarise(Pounds = sum(Pounds, na.rm = T))
```

##group into geographic units.  

Gulf of Maine = Maine, New Hampshire, Massachusetts 
Southern New England = Rhode Island, Connecticut, New York 
Mid Atlantic Bight = New Jersey, Deleware, Maryland, Virginia, North Carolina 
```{r}
landings$SUBREGION2 <- ifelse(landings$State == "MAINE" | landings$State == "NEW HAMPSHIRE" | landings$State == "MASSACHUSETTS", "Gulf of Maine", ifelse(landings$State == "RHODE ISLAND" | landings$State == "CONNECTICUT" | landings$State == "NEW YORK", "Southern New England", "Mid Atlantic Bight"))

landings <- landings %>% dplyr::select(-State) %>% group_by(Year, NMFS.Name, SUBREGION2) %>% summarise(Pounds = sum(Pounds, na.rm = T))
```

##add year lag
```{r}
landings$Year_to_use <- landings$Year + 1
```

##make new species column
```{r}
landings <- landings %>% separate(NMFS.Name, c("A", "B", "C"), ",")
cols <- c("A", "B", "C")
landings_2 <- landings %>% dplyr::select("A", "B", "C")

landings_2$spp <- paste(landings$C, landings$B, landings$A)

landings_2$spp <- gsub("NA", "", landings_2$spp)

landings_2$spp  <- trimws(landings_2$spp , which = c("both"))

landings <- cbind(landings, landings_2$spp)

landings$spp <- landings$...8

landings <- landings %>% dplyr::select(Year, SUBREGION2, Pounds, Year_to_use, spp)
```
##get out species to sleect

```{r}
#load data 
load('season2/XRA_6/output.rdata')
spp_class <- read.csv("spp_by_year_selected.csv")

out <- XRA_6
specnames_use <- colnames(out[["inputs"]][["y"]])

specnames_2 <- gsub("[[:punct:]]", " ", specnames_use) 

landings <- landings[landings$spp %in% specnames_2,]

landings <- left_join(landings, spp_class, by = c("spp" ="COMNAME"))
```

##create three variables, 
pelgagic fishing pressure, demersal fishing pressure, benthic fishing pressure 

```{r}
grouped_landings_2 <- landings %>% dplyr::select(SUBREGION2, Pounds, Year_to_use, depth_profile_pelagic_demersal, spp) %>% group_by(Year_to_use, SUBREGION2, depth_profile_pelagic_demersal) %>% summarise(Fishing_mortality = sum(Pounds, na.rm = T))

grouped_landings_2 <- grouped_landings_2 %>% spread(key = depth_profile_pelagic_demersal, value = Fishing_mortality)

colnames(grouped_landings_2) <- c("Year", "SUBREGION2", "F_benthic_sel", "F_demersal_sel", "F_pelagic_sel")
```

Okay so now we have demersal, pelagic, and benthic fishing pressure for each year and area. 

can add to xdata (season and annual) - note do this again if you want to use annual. 

season
```{r}
xdata_season <- read.csv("xdata_season.csv")
xdata_gridded <- read.csv("/Users/sarahroberts/Documents/ClimateOceanPlanning/R/gjam_time/xdata_gridded.csv")
xdata_gridded <- xdata_gridded %>% dplyr::select(groups, SUBREGION2)
mode <- function(x) { names(which.max(table(x))) }
region <- xdata_gridded %>% group_by(groups) %>% summarize(SUBREGION2 = mode(SUBREGION2))

xdata <- left_join(xdata_season, region, by = c("groups" = "groups"))

xdata <- xdata %>% dplyr::select(-F_benthic_sel, -F_demersal_sel, -F_pelagic_sel)
#xdata$SUBREGION2 <- ifelse(xdata$SUBREGION2 =="", "Gulf of Maine", xdata$SUBREGION2)
#xdata <- xdata %>% rename(times = time)
#colnames(xdata) <- gsub(pattern = "_", replacement = "", x = colnames(xdata))


xdata <- left_join(xdata, grouped_landings_2, by = c("year" = "Year", "SUBREGION2" = "SUBREGION2"))

write.csv(xdata, file = "xdata_season.csv")
```

annual 
```{r}
xdata_an <- read.csv("xdata_an.csv")
#xdata <- left_join(xdata_an, region, by = c("groups" = "groups"))


xdata <- xdata_an %>% dplyr::select(-F_benthic_sel, -F_demersal_sel, -F_pelagic_sel)
#xdata$SUBREGION2 <- ifelse(xdata$SUBREGION2 =="", "Gulf of Maine", xdata$SUBREGION2)
#xdata <- xdata %>% rename(times = time)
colnames(xdata) <- gsub(pattern = "_", replacement = "", x = colnames(xdata))

xdata <- left_join(xdata, grouped_landings_2, by = c("year" = "Year", "SUBREGION2" = "SUBREGION2"))

write.csv(xdata, file = "xdata_an.csv")

```

#START HERE 
#seasonal 
##load data 
```{r}
xdata_season <- read.csv("xdata_season.csv")
ydata_season_bio <- read.csv("ydata_season_bio.csv")
```

##setup 
seasonal, most common plus 7, season as a predictor in beta (interaction term)
```{r}
#xdata
xdata <- xdata_season
xdata <- xdata %>% dplyr::select(-X.1, -SUBREGION2, -F_benthic, -F_demersal, -F_pelagic)

#ydata 
spp_class <- read.csv("spp_by_year_selected.csv")
specnames <- as.vector(spp_class$SCINAME)
ydata <- ydata_season_bio
ydata[is.na(ydata)] <- 0

#select certain years 
ydata$year <- xdata$year

xdata <- subset(xdata, xdata$year > 1997)
ydata <- subset(ydata, ydata$year > 1997)

ydata <- ydata %>% dplyr::select(-year)
xdata$times <- xdata$times - (min(xdata$times)-1)

#hat if I just did ydata with a spec interaction 
ydata1 <- ydata[,colSums(ydata != 0) > 200] #at least 100 non-zero rows 
ydata1 <- ydata1[, colSums(ydata1 > 10) > 100] #at least 100 tows with 10 individuals caught 


#bay anchovy, jonah crab and atlantic rock crab , round herring, atlantic menhaden and striped anchovy are prey for a lot of species so I am going to add them back in. 

ydata1$ANCHOA.MITCHILLI <- ydata$ANCHOA.MITCHILLI
ydata1$CANCER.IRRORATUS<- ydata$CANCER.IRRORATUS
ydata1$CANCER.BOREALIS<- ydata$CANCER.BOREALIS
ydata1$ETRUMEUS.TERES <- ydata$ETRUMEUS.TERES
ydata1$BREVOORTIA.TYRANNUS <- ydata$BREVOORTIA.TYRANNUS
ydata1$ANCHOA.HEPSETUS <- ydata$ANCHOA.HEPSETUS
ydata1$ILLEX.ILLECEBROSUS <- ydata$ILLEX.ILLECEBROSUS

ydata <- ydata1


specnames <- colnames(ydata)
alphaSign <- read.csv("alphaSign_filled.csv")
alphaSign <- alphaSign %>% dplyr::select(-X)
rownames(alphaSign) <- colnames(alphaSign)
alphaSign <- alphaSign[,colnames(alphaSign) %in% specnames]
alphaSign <- alphaSign[rownames(alphaSign) %in% specnames,]
specnames <- colnames(alphaSign)
ydata <- ydata[,colnames(ydata) %in% specnames]
#lets change the column names in ydata and alphasign to be common names instead of scientific names 
spec <- as.data.frame(specnames)
spec$SCINAME <- spec$specnames
spp_selected <- read.csv("spp_by_year_selected.csv")
spp_selected_2 <- spp_selected %>% dplyr::select(SCINAME, COMNAME)
spp_selected_2$SCINAME <- gsub(pattern = ' ', replacement = ".", x = spp_selected_2$SCINAME)
spp_selected_2$COMNAME <- gsub(pattern = ' ', replacement = ".", x = spp_selected_2$COMNAME)
specnames_2 <- left_join(spec, spp_selected_2, by = "SCINAME")
specnames_2 <- as.vector(specnames_2$COMNAME)
colnames(ydata) <- specnames_2
colnames(alphaSign) <- specnames_2
rownames(alphaSign) <- specnames_2
rownames(alphaSign) <- colnames(alphaSign)
#create edata (when there are no species at all, effort is .001)
#create edata 
edata_bio <- ydata
edata_bio <- ifelse(rowSums(edata_bio>0), 1, .001)
edata_bio <- matrix(edata_bio, ncol = ncol(ydata), nrow = nrow(ydata))
colnames(edata_bio) <- colnames(ydata)
edata <- edata_bio


#alphaSign to just species I'm using now
alphaSign <- alphaSign[,colnames(alphaSign) %in% specnames_2]
alphaSign <- alphaSign[rownames(alphaSign) %in% specnames_2,]
specnames_2 <- colnames(alphaSign)
#fill missing times 
timeCol   <- 'times'
groupCol  <- 'groups'
groupVars <- c( 'groups' )
edata <- as.data.frame(edata) #get weird errors if I don't do this
ydata <- as.data.frame(ydata)
xdata <- as.data.frame(xdata)

tmp <- gjamFillMissingTimes(xdata, ydata, edata, groupCol, timeCol,
                            FILLMEANS = T, groupVars = groupVars,
                            typeNames = 'CA', missingEffort = .001) 

xdata <- tmp$xdata
ydata <- tmp$ydata
edata <- tmp$edata
edata <- ifelse(is.na(edata), .001,edata)
tlist <- tmp$timeList
snames   <- colnames(ydata)
trueE <- tmp$trueValues

#fixit below
colnames(xdata) <- c("insert","groups","times", "X", "year", "season", "groups2", "depth","times2", "month", "NAOJFM", "AMOyear", "SST", "SSAL", "chla", "SEDSIZE", "day", "season1", "date", "F_benthic_sel", "F_demersal_sel", "F_pelagic_sel" )

effort <- list(columns = 1:length(snames), values = edata)

#Fill in NAs from xdata here. 

#weird but for some reason the season column is no longer correct 
xdata$season <- xdata$season_1

#Fill in NAs from xdata here. 

xdata_na <- xdata[is.na(xdata$year), ]
xdata_na$year <- 1997
xdata_na$season1 <- as.integer(2)
#join to the one that has 1997 in it!
xdata_na <- left_join(xdata_na, xdata_season, by = c("year" = "year", "season1" = "season_1", "groups2" = "groups"))

xdata_na <- xdata_na %>% dplyr::select(insert, groups, times.x, X.x, year, groups2, depth.y, times.y, month.x, NAOJFM.y, AMOyear.y, oisst_full, SSAL_woa_insit, chla.y, SEDSIZE.y, day.y, season1, date.y, F_benthic_sel.y, F_demersal_sel.y, F_pelagic_sel.y)

colnames(xdata_na) <- colnames(xdata)
xdata_test <- left_join(xdata, xdata_na, by = c("insert" = "insert", "groups" = "groups"))

#ugh there is probably a better way to do this in a for loop but im lazy now
xdata_test$season1.x <- ifelse(is.na(xdata_test$year.x), xdata_test$season1.y, xdata_test$season1.x)
xdata_test$depth.x <- ifelse(is.na(xdata_test$depth.x), xdata_test$depth.y, xdata_test$depth.x)
xdata_test$times2.x <- ifelse(is.na(xdata_test$times2.x), xdata_test$times2.y, xdata_test$times2.x)
xdata_test$NAOJFM.x <- ifelse(is.na(xdata_test$NAOJFM.x), xdata_test$NAOJFM.y, xdata_test$NAOJFM.x)
xdata_test$AMOyear.x <- ifelse(is.na(xdata_test$AMOyear.x), xdata_test$AMOyear.y, xdata_test$AMOyear.x)
xdata_test$SST.x <- ifelse(is.na(xdata_test$SST.x), xdata_test$SST.y, xdata_test$SST.x)
xdata_test$SSAL.x <- ifelse(is.na(xdata_test$SSAL.x), xdata_test$SSAL.y, xdata_test$SSAL.x)
xdata_test$chla.x <- ifelse(is.na(xdata_test$chla.x), xdata_test$chla.y, xdata_test$chla.x)
xdata_test$SEDSIZE.x <- ifelse(is.na(xdata_test$SEDSIZE.x), xdata_test$SEDSIZE.y, xdata_test$SEDSIZE.x)
xdata_test$F_benthic_sel.x <- ifelse(is.na(xdata_test$F_benthic_sel.x), xdata_test$F_benthic_sel.y, xdata_test$F_benthic_sel.x)
xdata_test$F_demersal_sel.x <- ifelse(is.na(xdata_test$F_demersal_sel.x), xdata_test$F_demersal_sel.y, xdata_test$F_demersal_sel.x)
xdata_test$F_pelagic_sel.x <- ifelse(is.na(xdata_test$F_pelagic_sel.x), xdata_test$F_pelagic_sel.y, xdata_test$F_pelagic_sel.x)

xdata <- xdata_test[,1:21]
colnames(xdata) <- c("insert","groups","times", "X", "year", "groups2", "depth","times2", "month", "NAOJFM", "AMOyear", "SST", "SSAL", "chla", "SEDSIZE", "day", "season1", "date", "Fbenthicsel", "Fdemersalsel", "Fpelagicsel" )


```


setup 2; with fishing pressure with more species. 
```{r}
#xdata
xdata <- xdata_season
xdata <- xdata %>% dplyr::select(-X.1, -SUBREGION2, -F_benthic, -F_demersal, -F_pelagic)

#ydata 
spp_class <- read.csv("spp_by_year_selected.csv")
specnames <- as.vector(spp_class$SCINAME)
ydata <- ydata_season_bio
ydata[is.na(ydata)] <- 0

#select certain years 
ydata$year <- xdata$year

xdata <- subset(xdata, xdata$year > 1997)
ydata <- subset(ydata, ydata$year > 1997)

ydata <- ydata %>% dplyr::select(-year)
xdata$times <- xdata$times - (min(xdata$times)-1)

#hat if I just did ydata with a spec interaction 
ydata1 <- ydata[,colSums(ydata != 0) > 200] #at least 100 non-zero rows 
ydata1 <- ydata1[, colSums(ydata1 > 10) > 100] #at least 100 tows with 10 individuals caught 


#bay anchovy, jonah crab and atlantic rock crab , round herring, atlantic menhaden and striped anchovy are prey for a lot of species so I am going to add them back in. 

ydata1$ANCHOA.MITCHILLI <- ydata$ANCHOA.MITCHILLI
ydata1$CANCER.IRRORATUS<- ydata$CANCER.IRRORATUS
ydata1$CANCER.BOREALIS<- ydata$CANCER.BOREALIS
ydata1$ETRUMEUS.TERES <- ydata$ETRUMEUS.TERES
ydata1$BREVOORTIA.TYRANNUS <- ydata$BREVOORTIA.TYRANNUS
ydata1$ANCHOA.HEPSETUS <- ydata$ANCHOA.HEPSETUS
ydata1$ILLEX.ILLECEBROSUS <- ydata$ILLEX.ILLECEBROSUS

ydata <- ydata1


specnames <- colnames(ydata)
alphaSign <- read.csv("alphaSign_filled.csv")
alphaSign <- alphaSign %>% dplyr::select(-X)
rownames(alphaSign) <- colnames(alphaSign)
alphaSign <- alphaSign[,colnames(alphaSign) %in% specnames]
alphaSign <- alphaSign[rownames(alphaSign) %in% specnames,]
specnames <- colnames(alphaSign)
ydata <- ydata[,colnames(ydata) %in% specnames]
#lets change the column names in ydata and alphasign to be common names instead of scientific names 
spec <- as.data.frame(specnames)
spec$SCINAME <- spec$specnames
spp_selected <- read.csv("spp_by_year_selected.csv")
spp_selected_2 <- spp_selected %>% dplyr::select(SCINAME, COMNAME)
spp_selected_2$SCINAME <- gsub(pattern = ' ', replacement = ".", x = spp_selected_2$SCINAME)
spp_selected_2$COMNAME <- gsub(pattern = ' ', replacement = ".", x = spp_selected_2$COMNAME)
specnames_2 <- left_join(spec, spp_selected_2, by = "SCINAME")
specnames_2 <- as.vector(specnames_2$COMNAME)
colnames(ydata) <- specnames_2
colnames(alphaSign) <- specnames_2
rownames(alphaSign) <- specnames_2
rownames(alphaSign) <- colnames(alphaSign)
#create edata (when there are no species at all, effort is .001)
#create edata 
edata_bio <- ydata
edata_bio <- ifelse(rowSums(edata_bio>0), 1, .001)
edata_bio <- matrix(edata_bio, ncol = ncol(ydata), nrow = nrow(ydata))
colnames(edata_bio) <- colnames(ydata)
edata <- edata_bio


#alphaSign to just species I'm using now
alphaSign <- alphaSign[,colnames(alphaSign) %in% specnames_2]
alphaSign <- alphaSign[rownames(alphaSign) %in% specnames_2,]
specnames_2 <- colnames(alphaSign)
#fill missing times 
timeCol   <- 'times'
groupCol  <- 'groups'
groupVars <- c( 'groups' )
edata <- as.data.frame(edata) #get weird errors if I don't do this
ydata <- as.data.frame(ydata)
xdata <- as.data.frame(xdata)

tmp <- gjamFillMissingTimes(xdata, ydata, edata, groupCol, timeCol,
                            FILLMEANS = T, groupVars = groupVars,
                            typeNames = 'CA', missingEffort = .001) 

xdata <- tmp$xdata
ydata <- tmp$ydata
edata <- tmp$edata
edata <- ifelse(is.na(edata), .001,edata)
tlist <- tmp$timeList
snames   <- colnames(ydata)
trueE <- tmp$trueValues

#fixit below
colnames(xdata) <- c("insert","groups","times", "X", "year", "season", "groups2", "depth","times2", "month", "NAOJFM", "AMOyear", "SST", "SSAL", "chla", "SEDSIZE", "day", "season1", "date", "F_benthic_sel", "F_demersal_sel", "F_pelagic_sel" )

effort <- list(columns = 1:length(snames), values = edata)

#Fill in NAs from xdata here. 

#weird but for some reason the season column is no longer correct 
xdata$season <- xdata$season_1

#Fill in NAs from xdata here. 

xdata_na <- xdata[is.na(xdata$year), ]
xdata_na$year <- 1997
xdata_na$season1 <- as.integer(2)
#join to the one that has 1997 in it!
xdata_na <- left_join(xdata_na, xdata_season, by = c("year" = "year", "season1" = "season_1", "groups2" = "groups"))

xdata_na <- xdata_na %>% dplyr::select(insert, groups, times.x, X.x, year, groups2, depth.y, times.y, month.x, NAOJFM.y, AMOyear.y, oisst_full, SSAL_woa_insit, chla.y, SEDSIZE.y, day.y, season1, date.y, F_benthic_sel.y, F_demersal_sel.y, F_pelagic_sel.y)

colnames(xdata_na) <- colnames(xdata)
xdata_test <- left_join(xdata, xdata_na, by = c("insert" = "insert", "groups" = "groups"))

#ugh there is probably a better way to do this in a for loop but im lazy now
xdata_test$season1.x <- ifelse(is.na(xdata_test$year.x), xdata_test$season1.y, xdata_test$season1.x)
xdata_test$depth.x <- ifelse(is.na(xdata_test$depth.x), xdata_test$depth.y, xdata_test$depth.x)
xdata_test$times2.x <- ifelse(is.na(xdata_test$times2.x), xdata_test$times2.y, xdata_test$times2.x)
xdata_test$NAOJFM.x <- ifelse(is.na(xdata_test$NAOJFM.x), xdata_test$NAOJFM.y, xdata_test$NAOJFM.x)
xdata_test$AMOyear.x <- ifelse(is.na(xdata_test$AMOyear.x), xdata_test$AMOyear.y, xdata_test$AMOyear.x)
xdata_test$SST.x <- ifelse(is.na(xdata_test$SST.x), xdata_test$SST.y, xdata_test$SST.x)
xdata_test$SSAL.x <- ifelse(is.na(xdata_test$SSAL.x), xdata_test$SSAL.y, xdata_test$SSAL.x)
xdata_test$chla.x <- ifelse(is.na(xdata_test$chla.x), xdata_test$chla.y, xdata_test$chla.x)
xdata_test$SEDSIZE.x <- ifelse(is.na(xdata_test$SEDSIZE.x), xdata_test$SEDSIZE.y, xdata_test$SEDSIZE.x)
xdata_test$F_benthic_sel.x <- ifelse(is.na(xdata_test$F_benthic_sel.x), xdata_test$F_benthic_sel.y, xdata_test$F_benthic_sel.x)
xdata_test$F_demersal_sel.x <- ifelse(is.na(xdata_test$F_demersal_sel.x), xdata_test$F_demersal_sel.y, xdata_test$F_demersal_sel.x)
xdata_test$F_pelagic_sel.x <- ifelse(is.na(xdata_test$F_pelagic_sel.x), xdata_test$F_pelagic_sel.y, xdata_test$F_pelagic_sel.x)

xdata <- xdata_test[,1:21]
colnames(xdata) <- c("insert","groups","times", "X", "year", "groups2", "depth","times2", "month", "NAOJFM", "AMOyear", "SST", "SSAL", "chla", "SEDSIZE", "day", "season1", "date", "Fbenthicsel", "Fdemersalsel", "Fpelagicsel" )


```


##XRA_1 
fishing pressure in beta 
season in rho and betas as interaction with SST and depth 
```{r}
#set priors
xdata$season <- as.factor(xdata$season1)
#set priors
xdata$season <- as.factor(xdata$season)
xdata$Ftotal <- log(xdata$Fbenthicsel + xdata$Fdemersalsel + xdata$Fpelagicsel)
xdata$Fbenthicsel1 <- log(xdata$Fbenthicsel)
xdata$Fpelagicsel1 <- log(xdata$Fpelagicsel)
xdata$Fdemersalsel1 <- log(xdata$Fdemersalsel)


# effect on species growth.
rhoPrior <- list(lo = list(intercept = -1, SST = -2, SSAL = -2, chla = -2, season = -2), hi = list(intercept = 2,SST = 2, SSAL = 2, chla = 2, season = 2) )


#effect on e/imigration
betaPrior <- list(lo = list(intercept = -2, SEDSIZE = -2, depth = -2, SST = -2, SSAL = -2, chla = -2, season = -2, Fpelagicsel1 = -2, Fbenthicsel1 = -2, Fdemersalsel1 = -2), hi = list(intercept = 2, SEDSIZE = 2, depth = 2, SST = 2, SSAL = 2, chla = 2, season = 2, Fpelagicsel1 = 2, Fbenthicsel1 = 2, Fdemersalsel1 = 2) )

#prior lists  
priorList <- list( formulaBeta = ~ depth + SEDSIZE + SST + SSAL + chla  + chla*depth + SST*depth + season*SST + season*depth + Fbenthicsel1 + Fpelagicsel1 + Fdemersalsel1, formulaRho = as.formula(~  SST + SSAL + chla + season),  
                   betaPrior = betaPrior, rhoPrior = rhoPrior, alphaSign = alphaSign)


tmp <- gjamTimePrior( xdata, ydata, edata, priorList )

timeList <- mergeList(tlist, tmp)

#set limits on alpha priors
lo <- timeList$alphaPrior$lo
lo[lo < -4] <- -4
timeList$alphaPrior$lo <- lo 

hi <- timeList$alphaPrior$hi
hi[hi > 4 ] <- 4
timeList$alphaPrior$hi <- hi 

modelList <- list(typeNames = 'CA', ng = 5000, burnin = 1000,  
           timeList = timeList, effort = effort) 

#output
XRA_1 <- .gjam(~  depth + SEDSIZE + SST + SSAL + chla  + chla*depth + SST*depth + season*SST + season*depth + Fbenthicsel1 + Fpelagicsel1 + Fdemersalsel1, xdata=xdata, ydata=ydata, modelList=modelList, verbose=TRUE)
save(XRA_1, file='Fishing_pressure/seasonal/XRA_1/output.rdata')

plotPars <- list(PLOTALLY=T, SAVEPLOTS = T, SMALLPLOTS = T, GRIDPLOTS = T, outFolder = 'Fishing_pressure/seasonal/XRA_1')
gjamPlot(XRA_1, plotPars)

```
##XRA_2 
fishing pressure in beta as interaction 
season in rho and betas as interaction with SST and depth 
```{r}
#set priors
xdata$season <- as.factor(xdata$season1)


# effect on species growth.
rhoPrior <- list(lo = list(intercept = -1, SST = -2, SSAL = -2, chla = -2, season = -2), hi = list(intercept = 2,SST = 2, SSAL = 2, chla = 2, season = 2) )


#effect on e/imigration
betaPrior <- list(lo = list(intercept = -2, SEDSIZE = -2, depth = -2, SST = -2, SSAL = -2, chla = -2, season = -2, Fpelagicsel1 = -2, Fbenthicsel1 = -2, Fdemersalsel1 = -2), hi = list(intercept = 2, SEDSIZE = 2, depth = 2, SST = 2, SSAL = 2, chla = 2, season = 2, Fpelagicsel1 = 2, Fbenthicsel1 = 2, Fdemersalsel1 = 2) )

#prior lists  
priorList <- list( formulaBeta = ~ depth + SEDSIZE + SST + SSAL + chla  + chla*depth + SST*depth + season*SST + season*depth + SST*Fbenthicsel1 + SST*Fpelagicsel1 + SST*Fdemersalsel1, formulaRho = as.formula(~  SST + SSAL + chla + season),  
                   betaPrior = betaPrior, rhoPrior = rhoPrior, alphaSign = alphaSign)


tmp <- gjamTimePrior( xdata, ydata, edata, priorList )

timeList <- mergeList(tlist, tmp)

#set limits on alpha priors
lo <- timeList$alphaPrior$lo
lo[lo < -4] <- -4
timeList$alphaPrior$lo <- lo 

hi <- timeList$alphaPrior$hi
hi[hi > 4 ] <- 4
timeList$alphaPrior$hi <- hi 

modelList <- list(typeNames = 'CA', ng = 5000, burnin = 1000,  
           timeList = timeList, effort = effort) 

#output
XRA_2 <- .gjam(~  depth + SEDSIZE + SST + SSAL + chla  + chla*depth + SST*depth + season*SST + season*depth + SST*Fbenthicsel1 + SST*Fpelagicsel1 + SST*Fdemersalsel1, xdata=xdata, ydata=ydata, modelList=modelList, verbose=TRUE)
save(XRA_2, file='Fishing_pressure/seasonal/XRA_2/output.rdata')

plotPars <- list(PLOTALLY=T, SAVEPLOTS = T, SMALLPLOTS = T, GRIDPLOTS = T, outFolder = 'Fishing_pressure/seasonal/XRA_2')
gjamPlot(XRA_2, plotPars)

```
##XRA_3 
fishing pressure in rho as interaction term 
season in rho and betas as interaction with SST and depth 
get recipricoal condition number error here. Lets try F_total 
it only works with logging F (either total or separated out in to demersal etc)
```{r}
#set priors
xdata$season <- as.factor(xdata$season)
xdata$Ftotal <- log(xdata$Fbenthicsel + xdata$Fdemersalsel + xdata$Fpelagicsel)
xdata$Fbenthicsel1 <- log(xdata$Fbenthicsel)
xdata$Fpelagicsel1 <- log(xdata$Fpelagicsel)
xdata$Fdemersalsel1 <- log(xdata$Fdemersalsel)

# effect on species growth.
rhoPrior <- list(lo = list(intercept = -1, SST = -2, SSAL = -2, chla = -2, season = -2, Fpelagicsel1 = -2, Fbenthicsel1 = -2, Fdemersalsel1 = -2), hi = list(intercept = 2,SST = 2, SSAL = 2, chla = 2, season = 2, Fpelagicsel1 = 2, Fbenthicsel1 = 2, Fdemersalsel1 = 2) )

# rhoPrior <- list(lo = list(intercept = -1, SST = -2, SSAL = -2, chla = -2, season = -2, Ftotal = -2), hi = list(intercept = 2,SST = 2, SSAL = 2, chla = 2, season = 2, Ftotal = 2) )


#effect on e/imigration
betaPrior <- list(lo = list(intercept = -2, SEDSIZE = -2, depth = -2, SST = -2, SSAL = -2, chla = -2, season = -2), hi = list(intercept = 2, SEDSIZE = 2, depth = 2, SST = 2, SSAL = 2, chla = 2, season = 2) )

#prior lists  
priorList <- list( formulaBeta = ~ depth + SEDSIZE + SST + SSAL + chla  + chla*depth + SST*depth + season*SST + season*depth, formulaRho = as.formula(~  SST + SSAL + chla + season + Fpelagicsel1 + Fdemersalsel1 + Fbenthicsel1),  
                   betaPrior = betaPrior, rhoPrior = rhoPrior, alphaSign = alphaSign)


tmp <- gjamTimePrior( xdata, ydata, edata, priorList )

timeList <- mergeList(tlist, tmp)

#set limits on alpha priors
lo <- timeList$alphaPrior$lo
lo[lo < -4] <- -4
timeList$alphaPrior$lo <- lo 

hi <- timeList$alphaPrior$hi
hi[hi > 4 ] <- 4
timeList$alphaPrior$hi <- hi 

modelList <- list(typeNames = 'CA', ng = 800, burnin = 100,  
           timeList = timeList, effort = effort) 

#output
XRA_3 <- .gjam(~  depth + SEDSIZE + SST + SSAL + chla  + chla*depth + SST*depth + season*SST + season*depth, xdata=xdata, ydata=ydata, modelList=modelList, verbose=TRUE)
save(XRA_3, file='Fishing_pressure/seasonal/XRA_3/output.rdata')

plotPars <- list(PLOTALLY=T, SAVEPLOTS = T, SMALLPLOTS = T, GRIDPLOTS = T, outFolder = 'Fishing_pressure/seasonal/XRA_3')
gjamPlot(XRA_3, plotPars)

```

##XRA_4
fishing pressure in rho as interaction with SST 
season in rho and betas as interaction with SST and depth 
```{r}
#set priors
# effect on species growth.
rhoPrior <- list(lo = list(intercept = -1, SST = -2, SSAL = -2, chla = -2, season = -2, Fpelagicsel1 = -2, Fbenthicsel1 = -2, Fdemersalsel1 = -2), hi = list(intercept = 2,SST = 2, SSAL = 2, chla = 2, season = 2, Fpelagicsel1 = 2, Fbenthicsel1 = 2, Fdemersalsel1 = 2) )


#effect on e/imigration
betaPrior <- list(lo = list(intercept = -2, SEDSIZE = -2, depth = -2, SST = -2, SSAL = -2, chla = -2, season = -2), hi = list(intercept = 2, SEDSIZE = 2, depth = 2, SST = 2, SSAL = 2, chla = 2, season = 2) )

#prior lists  
priorList <- list( formulaBeta = ~ depth + SEDSIZE + SST + SSAL + chla  + chla*depth + SST*depth + season*SST + season*depth, formulaRho = as.formula(~  SST + + SSAL + chla + season + SST*Fpelagicsel1 + SST*Fdemersalsel1 + SST*Fbenthicsel1),  
                   betaPrior = betaPrior, rhoPrior = rhoPrior, alphaSign = alphaSign)


tmp <- gjamTimePrior( xdata, ydata, edata, priorList )

timeList <- mergeList(tlist, tmp)

#set limits on alpha priors
lo <- timeList$alphaPrior$lo
lo[lo < -4] <- -4
timeList$alphaPrior$lo <- lo 

hi <- timeList$alphaPrior$hi
hi[hi > 4 ] <- 4
timeList$alphaPrior$hi <- hi 

modelList <- list(typeNames = 'CA', ng = 800, burnin = 100,  
           timeList = timeList, effort = effort) 

#output
XRA_4 <- .gjam(~  depth + SEDSIZE + SST + SSAL + chla  + chla*depth + SST*depth + season*SST + season*depth, xdata=xdata, ydata=ydata, modelList=modelList, verbose=TRUE)
save(XRA_4, file='Fishing_pressure/seasonal/XRA_4/output.rdata')

plotPars <- list(PLOTALLY=T, SAVEPLOTS = T, SMALLPLOTS = T, GRIDPLOTS = T, outFolder = 'Fishing_pressure/seasonal/XRA_4')
gjamPlot(XRA_4, plotPars)

```

##XRA_5
fishing pressure in beta and rho (no interaction)
```{r}
#set priors
# effect on species growth.
rhoPrior <- list(lo = list(intercept = -1, SST = -2, SSAL = -2, chla = -2, season = -2, Fpelagicsel1 = -2, Fbenthicsel1 = -2, Fdemersalsel1 = -2), hi = list(intercept = 2,SST = 2, SSAL = 2, chla = 2, season = 2, Fpelagicsel1 = 2, Fbenthicsel1 = 2, Fdemersalsel1 = 2) )


#effect on e/imigration
betaPrior <- list(lo = list(intercept = -2, SEDSIZE = -2, depth = -2, SST = -2, SSAL = -2, chla = -2, season = -2, Fpelagicsel1 = -2, Fbenthicsel1 = -2, Fdemersalsel1 = -2), hi = list(intercept = 2, SEDSIZE = 2, depth = 2, SST = 2, SSAL = 2, chla = 2, season = 2, Fpelagicsel1 = 2, Fbenthicsel1 = 2, Fdemersalsel1 = 2) )

#prior lists  
priorList <- list( formulaBeta = ~ depth + SEDSIZE + SST + SSAL + chla  + chla*depth + SST*depth + season*SST + season*depth, formulaRho = as.formula(~  SST + SSAL + chla + season + Fpelagicsel1 + Fdemersalsel1 + Fbenthicsel1),  
                   betaPrior = betaPrior, rhoPrior = rhoPrior, alphaSign = alphaSign)


tmp <- gjamTimePrior( xdata, ydata, edata, priorList )

timeList <- mergeList(tlist, tmp)

#set limits on alpha priors
lo <- timeList$alphaPrior$lo
lo[lo < -4] <- -4
timeList$alphaPrior$lo <- lo 

hi <- timeList$alphaPrior$hi
hi[hi > 4 ] <- 4
timeList$alphaPrior$hi <- hi 

modelList <- list(typeNames = 'CA', ng = 800, burnin = 100,  
           timeList = timeList, effort = effort) 

#output
XRA_5 <- .gjam(~  depth + SEDSIZE + SST + SSAL + chla  + chla*depth + SST*depth + season*SST + season*depth + Fpelagicsel1 + Fdemersalsel1 + Fbenthicsel1, xdata=xdata, ydata=ydata, modelList=modelList, verbose=TRUE)
save(XRA_5, file='Fishing_pressure/seasonal/XRA_5/output.rdata')

plotPars <- list(PLOTALLY=T, SAVEPLOTS = T, SMALLPLOTS = T, GRIDPLOTS = T, outFolder = 'Fishing_pressure/seasonal/XRA_5')
gjamPlot(XRA_5, plotPars)

```

##XRA_6
fishing pressure in beta and rho (interaction with SST)
```{r}
#set priors
# effect on species growth.
rhoPrior <- list(lo = list(intercept = -1, SST = -2, SSAL = -2, chla = -2, season = -2, Fpelagicsel1 = -2, Fbenthicsel1 = -2, Fdemersalsel1 = -2), hi = list(intercept = 2,SST = 2, SSAL = 2, chla = 2, season = 2, Fpelagicsel1 = 2, Fbenthicsel1 = 2, Fdemersalsel1 = 2) )


#effect on e/imigration
betaPrior <- list(lo = list(intercept = -2, SEDSIZE = -2, depth = -2, SST = -2, SSAL = -2, chla = -2, season = -2, Fpelagicsel1 = -2, Fbenthicsel1 = -2, Fdemersalsel1 = -2), hi = list(intercept = 2, SEDSIZE = 2, depth = 2, SST = 2, SSAL = 2, chla = 2, season = 2, Fpelagicsel1 = 2, Fbenthicsel1 = 2, Fdemersalsel1 = 2) )

#prior lists  
priorList <- list( formulaBeta = ~ depth + SEDSIZE + SST + SSAL + chla  + chla*depth + SST*depth + season*SST + season*depth, formulaRho = as.formula(~  SST + SSAL + chla + season + Fpelagicsel1 + Fdemersalsel1 + Fbenthicsel1),  
                   betaPrior = betaPrior, rhoPrior = rhoPrior, alphaSign = alphaSign)


tmp <- gjamTimePrior( xdata, ydata, edata, priorList )

timeList <- mergeList(tlist, tmp)

#set limits on alpha priors
lo <- timeList$alphaPrior$lo
lo[lo < -4] <- -4
timeList$alphaPrior$lo <- lo 

hi <- timeList$alphaPrior$hi
hi[hi > 4 ] <- 4
timeList$alphaPrior$hi <- hi 

modelList <- list(typeNames = 'CA', ng = 800, burnin = 100,  
           timeList = timeList, effort = effort) 

#output
XRA_6 <- .gjam(~  depth + SEDSIZE + SST + SSAL + chla  + chla*depth + SST*depth + season*SST + season*depth + Fpelagicsel1 + Fdemersalsel1 + Fbenthicsel1, xdata=xdata, ydata=ydata, modelList=modelList, verbose=TRUE)
save(XRA_4, file='Fishing_pressure/seasonal/XRA_6/output.rdata')

plotPars <- list(PLOTALLY=T, SAVEPLOTS = T, SMALLPLOTS = T, GRIDPLOTS = T, outFolder = 'Fishing_pressure/seasonal/XRA_6')
gjamPlot(XRA_6, plotPars)

```
